---
title: "laboratory3"
author: "Zach F"
date: "1/31/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(here)
library(rio)
library(tidyverse)
library(gghighlight)
library(colorBlindness)
library(ggrepel)
library(scales)
library(ggthemes)
```

```{r}
honey <- import(here("data", "honeyproduct1.csv"))
```

## Number 1
Visualize the total production of honey across years by state Use color to highlight the west coast (Washington, Oregon, and California) with a different color used for each west coast state.

```{r}
honey_region <- honey %>% 
  mutate(region = ifelse(state %in% c("OR", "CA", "WA"), "West Coast", "Other"))

plot <- ggplot(honey_region, aes(year, totalprod)) +
  geom_point(color = "grey80") +
  geom_point(data = filter(honey_region, region == "West Coast"),
             aes(color = "West Coast")) +
  scale_color_manual(values = "#C55644") +
  theme(
    legend.title = element_blank(),
    legend.position = c(0.3, 0.9),
    legend.box.background = element_rect(color = "black"),
    legend.box.margin = margin(t = -.05, unit = "cm"),
  ) +
  labs(
    title = "States Total Honey Production Over Time",
    subtitle = "<span style = 'color: #C55644'>**West Coast**</span> compared to other states",
    caption = "Non West Coast states are shown in grey:
    West Coast states include California, Oregon and Washington",
    y = "Total Honey Production"
  ) +
  theme(plot.subtitle = ggtext::element_markdown())
plot
```


## Number 2
Reproduce the plot according three different kinds of color blindness, as well as a desaturated version.
```{r}
cvdPlot(plot)
```

## Number 3
Reproduce the plot using a color blind safe palette.
```{r}
plot_plain <- ggplot(honey_region, aes(year, totalprod)) +
  geom_point(color = "grey80") +
  geom_point(data = filter(honey_region, region == "West Coast"),
             aes(color = "West Coast")) +
  scale_color_manual(values = "#C55644") +
  theme(
    legend.title = element_blank(),
    legend.position = c(0.3, 0.9),
    legend.box.background = element_rect(color = "black"),
    legend.box.margin = margin(t = -.05, unit = "cm"),
  ) +
  labs(
    title = "States Total Honey Production Over Time",
    subtitle = "West Coast compared to other states",
    caption = "Non West Coast states are shown in grey",
    y = "Total Honey Production"
  ) #needed to remove the html code so that it would still read the same
plot_plain

plot_clbnd <- plot_plain +
  theme_igray() +
  scale_colour_colorblind()
plot_clbnd
```


